---
title: "Shiny Markdown"
author: "Clemens"
date: "August 3, 2018"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

You can use the "input" functions to add interactive compo-
nents to the document:

```{r}
library(shiny)

textInput("name", "What is your name?")
numericInput("age", "How old are you?", NA, min = 0, max = 150)
```


You can then refer to the values with input§name and input§age,
and the code that uses them will be automatically rerun whenever
they change.

I can't show you a live Shiny app here because Shiny interactions
occur on the server side. This means you can write interactive apps
without knowing JavaScript, but it means that you need a server to
run it on. This introduces a logistical issue: Shiny apps need a Shiny
server to be run online. When you run Shiny apps on your own
computer, Shiny automatically sets up a Shiny server for you, but
you need a public-facing Shiny server if you want to publish this
sort of interactivity online. That's the fundamental trade-off of
Shiny: you can do anything in a Shiny document that you can do in
R, but it requires someone to be running R.

Learn more about Shiny at http://shiny.rstudio.com/.