---
title: "Caching"
author: "Clemens"
date: "August 2, 2018"
output: html_document
---
Caching
Normally, each knit of a document starts from a completely clean
slate. This is great for reproducibility, because it ensures that you've
captured every important computation in code. However, it can be
painful if you have some computations that take a long time. The
solution is cache = TRUE. When set, this will save the output of the
chunk to a specially named file on disk. On subsequent runs, knitr
will check to see if the code has changed, and if it hasn't, it will reuse
the cached results.
The caching system must be used with care, because by default it is
based on the code only, not its dependencies. For example, here the
processed_data chunk depends on the raw_data chunk:

```{r raw_data}
rawdata <- readr::read_csv("a_very_large_file.csv")
```

```{r processed_data, cached = TRUE}
processed_data <- rawdata %>%
filter(!is.na(import_var)) %>%
mutate(new_variable = complicated_transformation(x, y, z))
```

Caching the processed_data chunk means that it will get rerun if
the dplyr pipeline is changed, but it won't get rerun if the
read_csv() call changes. You can avoid that problem with the
dependson chunk option:

```{r processed_data, cached = TRUE, dependson = "raw_data"}
processed_data <- rawdata %>%
filter(!is.na(import_var)) %>%
mutate(new_variable = complicated_transformation(x, y, z))
```

dependson should contain a character vector of every chunk that the
cached chunk depends on. knitr will update the results for the
cached chunk whenever it detects that one of its dependencies has
changed.

Note that the chunks won't update if a_very_large_fle.csv changes,
because knitr caching only tracks changes within the .Rmd file. If
you want to also track changes to that file you can use the
cache.extra option. This is an arbitrary R expression that will inva-
lidate the cache whenever it changes. A good function to use is
file.info(): it returns a bunch of information about the file
including when it was last modified. Then you can write:

```{r raw_data, cache.extra = file.info("a_very_large_file.csv")}
rawdata <- readr::read_csv("a_very_large_file.csv")
```

As your caching strategies get progressively more complicated, it's
a good idea to regularly clear out all your caches with
knitr::clean_cache().

I've used the advice of David Robinson to name these chunks: each
chunk is named after the primary object that it creates. This makes
it easier to understand the dependson specification.

