---
title: "R Markdown II"
author: "Clemens"
date: "August 2, 2018"
output: html_document
params:
  start: !r lubridate::ymd("2015-01-01")
  snapshot: !r lubridate::ymd_hms("2015-01-01 12:30:00")
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

As you can see, parameters are available within the code chunks as a
read-only list named params.
You can write atomic vectors directly into the YAML header. You
can also run arbitrary R expressions by prefacing the parameter
value with !r. This is a good way to specify date/time parameters.

In RStudio, you can click the "Knit with Parameters" option in the
Knit drop-down menu to set parameters, render, and preview the
report in a single user-friendly step.

Alternatively, if you need to produce many such parameterized
reports, you can call rmarkdown::render() with a list of params:

**R-Code**:
library(rmarkdown)
rmarkdown::render(
"fuel-economy.Rmd",
params = list(my_class = "suv"))


This is particularly powerful in conjunction with purrr:pwalk().
The following example creates a report for each value of class
found in mpg. First we create a data frame that has one row for each
class, giving the filename of report and the params it should be
given:

```{r}
library(tidyverse)
reports <- tibble(
class = unique(mpg$class),
filename = stringr::str_c("fuel-economy-", class, ".html"),
params = purrr::map(class, ~ list(my_class = .)))
reports
```

Then we match the column names to the argument names of ren
der(), and use purrr's parallel walk to call render() once for each
row:



**R-Code**:
reports %>%
select(output_file = filename, params) %>%
purrr::pwalk(rmarkdown::render, input = "fuel-economy.Rmd")


